Layered Formal Verification of an Industrial TCP/IP Stack

### 2021-05-21 SecDev21/paper.tex
cleanup intro

By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}. At the same time, it is made evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}. Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices. In 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used potentially by billions of IoT devices for more than 13 years \cite{zero_days_vuln:ARMIS_white_paper}, with some of them even allowing remote code execution.

@@

Existing research deals with formally verifying several protocols TCP/IP stack levels. For example, the work in \emph{miTLS} \cite{bhargavan2013implementing} formally verifies an SSL/TLS protocol implementation, and the work in \cite{Reiher_2020} uses a technology called RecordFlux to safely parse data segments. Given the reliance of the other software-level protocols on the TCP protocol, formalizing the its specification and verifying its implementation is critical. Achieving these two goals in practice is challenging. This is mainly because TCP's specification, \cite{rfc793}, is written in English rather than any formal language, leaving many parts underspecified. Moreover, the TCP protocol is inherently concurrent, making it complex to specify and verify. Previous work focused on formalizing the protocol to either prove key properties \cite{smith1996formal} or fully verify the TCP protocol's specifications \cite{ridge2008rigorous} using the HOL proof assistant tool. While for the later work, the authors claim that it is possible from their formalization to extract an actual implementation in Haskell, this has not been achieved yet. At this point of writing, we are not aware of any formally verified TCP protocol implementation.

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK \cite{mccormick_chapin_2015}. SPARK 2014 is a programming language designed as a subset of the Ada language to produce highly reliable software through formal verification. For this work, the CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company \cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack. By using as a starting point such a well-tested and widely used TCP/IP implementation, the IoT industry can have an immediate benefit by adopting this work's security-enhanced version of the CycloneTCP library.

>>>>

By 2030 it is estimated that one trillion Internet of Things (IoT) devices will be deployed and be part of every aspect of our daily life~\cite{IoT_route:ARM_white_paper}. At the same time, it is made evident that security for such systems is a major challenge~\cite{Sec_IoT_Challenges:paper,Cyber_Systems_security:paper}. Any vulnerability in software libraries that are widely used by embedded systems can potentially open the door for compromising the security of billions of deployed IoT devices. In 2019, critical security vulnerabilities were discovered in a TCP implementation that has been used by billions of IoT devices for more than 13 years~\cite{zero_days_vuln:ARMIS_white_paper}, some of them allowing remote code execution.

@@

Existing research deals with formally verifying several protocols of TCP/IP stack levels. Early work formalized the TCP protocol to prove some key properties of the protocol itself~\cite{smith1996formal}. More recently, researchers have formally specified TCP, UDP and the Sockets API in HOL~\cite{tcp2005sigcomm,ridge2008rigorous} and used the resulting specification for testing existing libraries. Others have formally specified and implemented the SSL/TLS protocol in F*\cite{bhargavan2013implementing}, and integrated the resulting implementation in existing libraries. To the best of our knowledge, no prior work attempted to formally verify an implementation of TCP, although the authors of~\cite{ridge2008rigorous} suggest that it would possible to extract an actual implementation in Haskell from their specification.

In this paper, we take a practical approach, where the security of an existing TCP/IP stack implementation written in ANSI C programming language is incrementally enhanced by replacing parts of its code with formally verified code in SPARK~\cite{mccormick_chapin_2015}, a subset of the Ada programming language supported by formal verification tools. We chose CycloneTCP, a professional-grade embedded TCP/IP library developed by the Oryx Embedded company~\cite{CycloneTCP}. \Cref{Fig:TcpStack} shows the protocols supported by the library. This work focuses on the TCP and socket components of the stack. The library's TCP implementation is meant to conform with the RFC 793 protocol specifications \cite{rfc793}. The quality of the CycloneTCP is acknowledged by the AMNESIA:33 report \cite{AMNESIA33}, which classifies it as one of the most resilient TCP/IP stack.
